Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-14 shanghai.tex
All the definitions are in now.

\begin{proof}
 To prove the proposition we exhibit a full and faithful monoidal functor $L_G : \QCf(G) \to \QCb(G)$.
 %
 We will do this by constructing a faithful monoidal functor $L : C_1(G) \to \QCb(G)$ 
 such that $L(\alpha,T)$ is an isomorphism precisely when $(h,T)$ is a weak isomorphism. 
 Factoring along $C_1(G) \to \QCb(G)$ will produce the full and faithful monoidal functor we seek.

@@

 Now consider the localization $C(G)[W^{-1}]$ of $C(G)$ at weak isomorphisms.
 Morphisms in $C(G)[W^{-1}]$ are simple to describe because category $C(G)$ 
 admits a calculus of left fractions\footnote{See {\tt http://ncatlab.org/nlab/show/calculus+of+fractions}.} for weak isomorphisms: 
 every isomorphism in $C(G)$ is a weak isomorphism; 
 push-outs exist in $C(G)$ and the push-out of a weak isomorphism is a weak isomorphism; 
 and weak isomorphisms enjoy the left-cancellation property. 
 It follows that a morphism in $C(G)[W^{-1}]$ is the equivalence class of 
  \begin{equation}\label{eq:morGCf}
   \begin{tikzcd}
   (f,\psi) \arrow{r}{(\alpha_1,T_1)} & (f_1,\psi_1) & \arrow[swap]{l}{(\alpha'_1,T'_1)} (f',\psi')
   \end{tikzcd}
  \end{equation}
 where $(\alpha',T') : (f,\psi) \to (f',T')$ is any map in $C(G)$ 
 and $(\alpha'_1,T'_1) : (f_1,\psi_1) \to (f',\psi')$ is a weak isomorphism,
 and therefore given by a diagram of the form

 Observe that $L$ takes weak isomorphisms in $C(G)$ to isomorphisms in $\Loc(G)$. 
 To see this, let $(h,T) : (f,\psi)\to (f',\psi')$ be a weak isomorphism, 
 where $f:H\to G$ and $f':H'\to G$ as above. 
 Since $T$ is an isomorphism, we may replace $V'$ with $V$ and $T$ with $\id_V$, without loss of generality. 
 Then the equivariance condition on $T$ becomes $\psi' = \psi\circ h$ on the finite etale group scheme $\ker f'$. 
 Since $\alpha : H'\to H$ is an etale cover, it admits a section $\alpha' : H \to H'$, 
 which provides the inverse to $L(\alpha,T) : L(f,\psi) \to L(f',\psi')$, above. 
 It is also clear that if $L(\alpha,T)$ is an isomorphism of local systems, 
 then $T$ must be an isomorphism of vector spaces, 
 whence $(\alpha,T)$ is a weak isomorphism: $W$ is saturated.

 Since $L : C(G) \to \Loc(G)$ takes weak isomorphisms to isomorphisms,
 it factors (uniquely!) through the localisation functor $C(G) \to C(G)[W^{-1}]$ 
 to define a faithful functor $L_{W} : C(G)[W^{-1}] \to \Loc(G)$. 
 Since $W$ is saturated, the functor $L_{W} : C(G)[W^{-1}] \to \Loc(G)$ is full.

@@

\begin{proof}
As in Lemma~\ref{lemma:finite-pull-back}, we have $m^*(f,\psi) = (f_m,\psi_m)$ where $f_m$ and $\psi_m$ are pictured below, where $H_m = H\times_G(G\times G)$.
\[
\begin{tikzcd}
V & \Aut(V) & \arrow[swap]{l}{\psi} A \arrow[hook]{r} & H \arrow{r}{f} &  G\\
  &  & \arrow[dashed]{ul}{\psi_g} \arrow[swap]{u}{(m_f)\vert_{\ker m_f}} \ker f_m \arrow[hook]{r} & \arrow[swap]{u}{m_f} H_m \arrow{r}{f_m} & \arrow{u}{m} G\times G.
\end{tikzcd}
\]
On the other hand, $(f,\psi) \boxtimes (f,\psi) = (f\times f, \psi\otimes \psi)$.
Since $f\circ m_H = m \circ (f\times f)$, by the universal property of pull-back, there is a unique $\theta : H\times H \to H_m$ making the following commute.
\[
\begin{tikzcd}
\ &  & \arrow{ld}{\psi} A \arrow[hook]{r} & H \arrow{r}{f} &  G\\
V \arrow{d}{U} & \Aut(V)  & \arrow{l}{\psi_g} \arrow{u} A_m \arrow[hook]{r} & \arrow[swap]{u}{m_f} H_m \arrow{r}{f_m} & \arrow{u}{m} G\times G\\
V\otimes V & \Aut(V\otimes V) & \arrow[swap]{l}{\psi\otimes\psi} \arrow[bend left]{uu}{m_H} \arrow[swap]{u}{\theta\vert_{A\times A}} A \times A \arrow[hook]{r}{} & \arrow[bend left]{uu}{m_H} \arrow[swap]{u}{\theta} H \times H \arrow{r}{f\times f} & \arrow[equal]{u} G\times G\\
\end{tikzcd}
\]
Choose a basis $\{ v_1 \}$ for $V$; then $\{ v_1\otimes v_1 \}$ is a basis for $V\otimes V$. Define $U : V \to V\otimes V$ by $U(v_1) = v_1\otimes v_1$. Then the lower part of this diagram defines a weak isomorphism $(\theta,U): m^*(f,\psi) \to (f,\psi)\boxtimes (f,\psi)$  in $C_1(G)$.

While $\theta$ was canonical, $U$ was not, as it depended on the choice of a basis for $V$. However, after passing from $C_1(G)$ to $\QCf(G)$, the resulting isomorphism $\mu(f,\psi): m^*(f,\psi) \to (f,\psi)\boxtimes (f,\psi)$ is canonical in $\QCf(G)$.
\end{proof}

@@

\'Etale group schemes form a counterpoint to connected groups, since the component group
is always \'etale.  The functor $G \mapsto G(\bFq)$ defines an equivalence of categories
between the category of \'etale group schemes over $\Fq$ and the category of groups equipped
with an action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on the group.

For such a $\Gal(\bFq/\Fq)$-group $G(\bFq)$, 
\Clifton{Re-phrase.}
we define a category of quasicharacter sheaves $\QC(G(\bFq))$.
An object in $\QC(G(\bFq))$ is given by the data of
\begin{enumerate}
 \labitem{(cs.0)}{cs.0} an indexed set of one-dimensional
  $\EE$-vector spaces $\gcs{L}_x$, as $x$ runs over
  $G(\bFq)$;

@@

 \labitem{(cs.2)}{cs.2} an indexed set of isomorphisms $\phi_{x} : \gcs{L}_{\Frob{\Fq}(x)} \to \gcs{L}_x$
  such that
  \[
   \begin{tikzcd}[row sep=40]
    \gcs{L}_{\Frob{\Fq}(x)+\Frob{\Fq}(y)} \arrow[swap]{d}{\phi_{x+y}} \arrow{rr}{\mu_{\Frob{\Fq}(x),\Frob{\Fq}(y)}}
    && \gcs{L}_{\Frob{\Fq}(x)}\otimes \gcs{L}_{\Frob{\Fq}(y)} \dar{\phi_x \tight{0}{\otimes}{0} \phi_y} \\
    \gcs{L}_{x+y} \arrow{rr}{\mu_{x,y}}
    && \gcs{L}_x \otimes\gcs{L}_y
   \end{tikzcd}
  \]
  commutes, for all $x,y\in G(\bFq)$.
\end{enumerate}
A morphism $\alpha : \cs{L} \to \cs{L'}$ in $\QC(G(\bFq))$ is given by 
\begin{enumerate}
 \labitem{(cs.3)}{cs.3} an indexed set $\alpha_x : \gcs{L}_x \to \gcs{L'}_x$
  of linear transformations (of one-dimensional $\EE$-vector spaces),
  as $x$ runs over $G(\bFq)$, such that
  \[
   \begin{tikzcd}[column sep=40]
    \arrow[swap]{d}{\phi_x} \gcs{L}_{\Frob{\Fq}(x)} \rar{\alpha_{\Frob{\Fq}(x)}} & \gcs{L'}_{\Frob{\Fq}(x)} \dar{\phi_x'}
    &\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}}& \arrow[swap]{d}{\mu} \gcs{L}_{x+y} \rar{\alpha_{x+y}} & \gcs{L'}_{x+y} \dar{\mu'_{x+y}} \\
    \gcs{L}_x \rar{\alpha_x} & \gcs{L'}_x
    & {} & \gcs{L}_x\otimes\gcs{L}_y \rar{\alpha_x\otimes \alpha_y} & \gcs{L'}_x \otimes \gcs{L'}_y
   \end{tikzcd}
  \]
  both commute.
\end{enumerate}

The functor $\QC(G) \to \QC(G(\bFq))$ defined by
$(\gcs{L}, \mu, \phi) \mapsto (\gcs{L}_x = x^*\gcs{L}, \mu_{x,y} = (x \times y)^* \mu, \phi_x = x^*\phi)$
is easily seen to be an equivalence.

We may now define a \cdef{global section}
of a quasicharacter sheaf $\cs{L}$ on an \'etale group scheme $G$ as a choice $s(x) \in \gcs{L}_x$ for every
$x \in G(\bFq)$, satisfying
 \begin{equation}\label{section}
  \forall x,y \in G(\bFq), \qquad \mu_{x,y}(s(x+y)) = s(x)\otimes s(y).
 \end{equation}
\Clifton{I've been putting lemmata after the propositions that use them, 
to keep the flow of the sections clear. 
Would you mind if I do the same in this section?}

\begin{lemma}\label{lemma:section}
 If $\cs{L}$ is a quasicharacter sheaf on a commutative, \'etale group scheme $G$
 with $G(\bFq)$ finitely generated then nontrivial global sections of $\cs{L}$ exist.
\end{lemma}

@@

These global sections will allow us to construct the isomorphism in the following proposition.

@@

\begin{proof}
  Suppose $\cs{L}$ is a quasicharacter sheaf on $G$.  Using Lemma~\ref{lemma:section}, we may choose a global section $s_{\cs{L}}$
  of $\cs{L}$.  We will construct a group isomorphism between $\QCiso{G(\bFq)}$ and
  $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$ using $s_{\cs{L}}$, then prove independence from the choice of section.

@@

  Although we used a global section $s_\cs{L}$ in the definition of
  $\tau_\cs{L}$, the class of $\tau_\cs{L}$ does not depend on this
  choice. To see why, let $s_1$ and $s_2$ be two global sections of
  $\cs{L}$ and let $\tau_1$ and $\tau_2$ be the cocycles defined by $s_1$ and $s_2$.
  For each $x \in G(\bFq)$ there is a unique scalar
  $a(x) \in \EEx$ such that $s_2(x) = a(x) s_1(x)$. Then \eqref{section} and
  condition~\ref{cs.1} together imply $a(x+y) = a(x)a(y)$, so
  $a\in G(\bFq)^*$.  The coboundary $\tau_0 : \Weil{\Fq} \to G(\bFq)^*$
  defined by $\tau_0(\Frob{\Fq}) = \,^\Frob{\Fq} a \cdot a^{-1}$ satisfies $\tau_2 = \tau_0 \tau_1$,
  so $\tau_2$ and $\tau_1$ are cohomologous. Thus
  \begin{equation}\label{cohomologous}
    \obj \QC(G(\bFq)) \longrightarrow \Hh^1(\Weil{\Fq},G(\bFq)^*),
    \qquad \text{defined by} \qquad \cs{L} \mapsto [\tau_\cs{L}],
  \end{equation}
  is independent of the choice of global sections made above.

@@

With Theorem~\ref{thm:snake} on hand, we can see why neither bounded quasicharacter sheaves nor finite quasicharacter sheaves are up to the task of geometrizing characters of $G(\Fq)$ in the generality we need.  To see this, consider the case when $G$ is the discrete etale group scheme $\ZZ$. Let $\chi : G(\Fq) \to \EEx$ be the character determined by $\chi(1) = \ell$ and let $\cs{L}$ be a quasicharacter sheaf in the isomorphism class corresponding to the character $\chi : G(\Fq) \to \EEx$ under Theorem~\ref{thm:snake}. Then $\trFrob{\cs{L}}(1) = \ell$. Suppose $\cs{L}$ was a bounded character sheaf. Then the restriction of $\cs{L}$ to the connected component $1\in G$ would be a rank-1 local system on $1$, thus an $\ell$-adic chacter $\rho_1 : \pi_1(1,{\bar 1}) \to \EEx$ with $\rho_1(\Frob{\Fq}) = \ell$. But $\pi_1(1,{\bar 1}) = \Gal(\bFq/\Fq) \iso {\hat \ZZ}$ is compact, while $\rho_1(\pi_1(1,{\bar 1}))$ is not compact (it contains $\ell^\ZZ$). Since the Weil sheaf $\gcs{L}$ does not descend to a local system on $G$, the geometric character $\cs{L}$ is not a bounded geometric character. We can play the same game with the character $\chi : G(\Fq) \to \EEx$ defined by $\chi(1) = 1+\ell$. In this case, the corresponding geometric character $\cs{L}$ is bounded but not finite, since $1+\ell$ is not a root of unity. This simple example illustrates why the categories $\QCb(G)$ and $\QCf(G)$ are inadequate to geometrize characters of $G(\Fq)$.

@@

Let $K$ be a non-Archimedean local field with residue field $\Fq$ 
and let $T$ be an algebraic torus over $K$.

@@

Let $K$ be a local field with residue field $\Fq$.
%(Note that we place no restriction on the characteristic of $K$.)
Let $T$ be an algebraic torus over $K$.

%In this section we introduce a group scheme $\GN{T}$ over $\Fq$ obtained by applying the Greenberg functor to the N\'eron lft-model for $T$ and establish some basic facts about $\GN{T}$

@@

The lft-N\'eron model for $T$, denoted here by $\TT$, 
is a separated, smooth, commutative group scheme locally of finite type over $\OK$ 
(the ring of integers of $K$)
with generic fibre $T$, that enjoys the N\'eron mapping property: if
$S$ is a smooth scheme over $\OK$ then every morphism $S_\eta \to T$
over $K$ extends uniquely to a morphism $S \to \TT$. Equivalently, the
N\'eron mapping property states that the function
$\Hom_{\OK} (S,\TT) \to \Hom_K(S_\eta,T)$, given by restriction to
generic fibres, is a bijection; 
in particular, taking $S = \Spec{\OK}$, this implies $\TT(\OK) = T(K)$.

For each positive integer $d$, let $\TT_{d}$ be the group scheme obtained from $\TT$ by base change along $\OK \to \OK/\pK^d$, where $\pK$ is the maximal ideal in $\OK$. 
Observe that $\OK/\pK^d$ is an Artin ring and $\TT_{d}$ is smooth and locally of finite type over $\OK/\pK^d$.
%
Let \cdef{$\GN{T}_{d} = \Gr(\TT_d)$} be the group scheme over $\Fq$ produced by applying
the Greenberg functor to $\TT_{d}$; this is the \cdef{Greenberg transform of $\TT_d$}.
See \cite{Greenberg:2} and \cite[Ch. 9, \S 6]{BLR} and the first few sections of \cite{Stasinski} for background on the Greenberg functor. The fundamental properties of $\GN{T}_d$ that we need are:
 \begin{enumerate}
   \labitem{(GN.0)}{dFq} $\GN{T}_d(\Fq) = \TT(\OK/\pK^d)$, canonically.
  \labitem{(GN.1)}{GNd} $\GN{T}_d$ is a smooth, commutative group scheme, locally of finite type over $\Fq$, and the component group of $\GN{T}_d$ is geometrically finitely generated.
  \end{enumerate}
In fact, $\pi_0(\GN{T}_d) = \pi_0(\TT) \times_{\Spec{\OK}} \Spec{\Fq}$, and the latter is finitely generated by \cite{Xarles}.

%\input{application}
\begin{theorem}\label{thm:application}
  Let $T$ be an algebraic torus over a non-Archimedean local field $K$.
  Let $\GN{T}_d$ be as above.  For each positive integer $d$, the
  Picard group of the category of quasicharacter sheaves on $\GN{T}_d$ is
  canonically isomorphic to the group of characters of $T(K)$ of depth
  less than $d$:
  \[
  \QCiso{\GN{T}_d} \iso \Hom_{< d}(T(K),\EEx).
  \]
\end{theorem}

\begin{proof}
  By \ref{GNd}, we may apply Theorem~\ref{thm:snake} to
  $\GN{T}_d$ which provides a canonical isomorphism
  $\QCiso{\GN{T}_d} \iso \GN{T}_d(\Fq)^*$.
  By \ref{dFq}, $\GN{T}_d(\Fq) \iso \TT(\OK/\pK^d)$ canonically.
  To finish the proof we recall that the filtration of $T(K)$ used to
  define depth has the property that $\TT(\OK/\pK^d) \cong T(K) / T(K)_d$
  \cite[Prop 5.2]{Yu}, so characters of $\GN{T}_d(\Fq)$ 
  are in bijection with characters of $T(K)$ vanishing on $T(K)_d$. 
  %Note that $d$ is a positive integer.  
\end{proof}

@@

%\input{pro}
We may also assemble the group $\Fq$-schemes $\TT_{d}$ into one
glorious group pro-scheme over $\Fq$:
\begin{equation}\label{projlim}
  \cdef{\GN{T}} \ceq \varprojlim_{d\in \ZZ_{>0}} \GN{T}_{d}.
\end{equation}
This is the \cdef{Greenberg transform of the N\'eron model of $T$}.
 Some fundamental properties of $\GN{T}$ are:
 \begin{enumerate}
   \labitem{(GN.2)}{T(K)} $\GN{T}(\Fq) = T(K)$, canonically;
  \labitem{(GN.3)}{scheme} $\GN{T}$ is a commutative group scheme over $\Fq$.
  \end{enumerate}
This gives us a natural way to geometrize all quasicharacters of $T(K)$
without reference to depth.
Let \cdef{$\QC(\GN{T})$} be the ind-category $\varinjlim_{d\in \ZZ_{>0}} \QC(\GN{T}_d)$;
we refer to objects of $\QC(\GN{T})$ as \cdef{quasicharacter sheaves on $T$}.

\begin{corollary}
Let $T$ be an algebraic torus over a non-Archimedean local field $K$.
The Picard group of the category $\QC(\GN{T})$ of quasicharacter sheaves on $T$ is canonically isomorphic to the group of quasicharacters of $T(K)$:
  \[
 \QCiso{\GN{T}} \iso \Hom_\text{ad}(T(K),\EEx).
  \]
\end{corollary}

>>>>

\begin{proof}
 To prove the proposition we exhibit a full and faithful monoidal functor $L_G : \QCf(G) \to \QCb(G)$.
 %
 We will do this by constructing a faithful monoidal functor $L : C_1(G) \to \QCb(G)$ 
 such that $L(\alpha,T)$ is an isomorphism precisely when $(\alpha,T)$ is a weak isomorphism. 
 Factoring along $C_1(G) \to \QCb(G)$ will produce the full and faithful monoidal functor we seek.

@@

 Now consider the localization $C(G)[W^{-1}]$ of $C(G)$ at weak isomorphisms.
 Morphisms in $C(G)[W^{-1}]$ are simple to describe because category $C(G)$ 
 admits a calculus of left fractions
 %\footnote{See {\tt http://ncatlab.org/nlab/show/calculus+of+fractions}.} 
 for weak isomorphisms:
% every isomorphism in $C(G)$ is a weak isomorphism; 
% push-outs exist in $C(G)$ and the push-out of a weak isomorphism is a weak isomorphism; 
% and weak isomorphisms enjoy the left-cancellation property. 
% It follows that
 a morphism in $C(G)[W^{-1}]$ is the equivalence class of 
  \begin{equation}\label{eq:morGCf}
   \begin{tikzcd}
   (f,\psi) \arrow{r}{(\alpha_1,T_1)} & (f_1,\psi_1) & \arrow[swap]{l}{(\beta_1,U_1)} (f',\psi')
   \end{tikzcd}
  \end{equation}
 where $(\alpha',T') : (f,\psi) \to (f',T')$ is any map in $C(G)$ 
 and $(\beta_1,U_1) : (f_1,\psi_1) \to (f',\psi')$ is a weak isomorphism.

 Observe that $L$ takes weak isomorphisms in $C(G)$ to isomorphisms in $\Loc(G)$. 
 To see this, suppose $(\beta,U) : (f,\psi)\to (f',\psi')$ is a weak isomorphism. 
 Since $U$ is an isomorphism, we may replace $V'$ with $V$ and $U$ with $\id_V$, without loss of generality. 
 Then the equivariance condition on $U$ becomes $\psi' = \psi\circ \beta$ on the finite etale group scheme $\ker f'$. 
 Since $f$ and $f'$ are finite etale, and $f' = f\circ \beta$, and $\beta$ is surjective, $\beta$ is an etale cover; therefore, $\beta$ admits a section $\beta' : H \to H'$. This provides the inverse to $L(\beta,U) : L(f,\psi) \to L(f',\psi')$.

 Conversely, if $L(\alpha,T)$ is an isomorphism of local systems, 
 then $T$ must be an isomorphism of vector spaces and $\beta$ must be surjective, whence $(\alpha,T)$ is a weak isomorphism: $W$ is saturated.

 Since $L : C(G) \to \Loc(G)$ takes weak isomorphisms to isomorphisms,
 it factors, uniquely, through the localisation functor $C(G) \to C(G)[W^{-1}]$ 
 to define a faithful functor $L_{W} : C(G)[W^{-1}] \to \Loc(G)$. 
 Since $W$ is saturated, the functor $L_{W} : C(G)[W^{-1}] \to \Loc(G)$ is full.

@@

\begin{proof}
As in Lemma~\ref{lemma:finite-pull-back}, we have $m^*(f,\psi) = (f_m,\psi_m)$ where $f_m$ and $\psi_m$ are pictured below, where $H_m = H\times_G(G\times G)$.
\[
\begin{tikzcd}
V & \Aut(V) & \arrow[swap]{l}{\psi} A \arrow[hook]{r} & H \arrow{r}{f} &  G\\
  &  & \arrow[dashed]{ul}{\psi_g} \arrow[swap]{u}{(m_f)\vert_{\ker m_f}} \ker f_m \arrow[hook]{r} & \arrow[swap]{u}{m_f} H_m \arrow{r}{f_m} & \arrow{u}{m} G\times G.
\end{tikzcd}
\]
On the other hand, $(f,\psi) \boxtimes (f,\psi) = (f\times f, \psi\otimes \psi)$.
Since $f\circ m_H = m \circ (f\times f)$, by the universal property of pull-back, there is a unique $\theta : H\times H \to H_m$ making the following commute.
\[
\begin{tikzcd}
\ &  & \arrow{ld}{\psi} A \arrow[hook]{r} & H \arrow{r}{f} &  G\\
V \arrow{d}{T} & \Aut(V)  & \arrow{l}{\psi_g} \arrow{u} A_m \arrow[hook]{r} & \arrow[swap]{u}{m_f} H_m \arrow{r}{f_m} & \arrow{u}{m} G\times G\\
V\otimes V & \Aut(V\otimes V) & \arrow[swap]{l}{\psi\otimes\psi} \arrow[bend left]{uu}{m_H} \arrow[swap]{u}{\theta\vert_{A\times A}} A \times A \arrow[hook]{r}{} & \arrow[bend left]{uu}{m_H} \arrow[swap]{u}{\theta} H \times H \arrow{r}{f\times f} & \arrow[equal]{u} G\times G\\
\end{tikzcd}
\]
Choose a basis $\{ v_1 \}$ for $V$; then $\{ v_1\otimes v_1 \}$ is a basis for $V\otimes V$. Define $T : V \to V\otimes V$ by $T(v_1) = v_1\otimes v_1$. Then the lower part of this diagram defines a weak isomorphism $(\theta,T): m^*(f,\psi) \to (f,\psi)\boxtimes (f,\psi)$  in $C_1(G)$.

While $\theta$ was canonical, $T$ was not, as it depended on the choice of a basis for $V$. However, after passing from $C_1(G)$ to $\QCf(G)$, the resulting isomorphism $\mu(f,\psi): m^*(f,\psi) \to (f,\psi)\boxtimes (f,\psi)$ is canonical in $\QCf(G)$.
\end{proof}

@@

\'Etale group schemes form a counterpoint to connected groups, since the component group of a group scheme locally of finite type over $\Fq$ is an \'etale group scheme (locally of finite type) over $\Fq$ \cite{vdG&M}*{III, \S 4}.  The functor $G \mapsto G(\bFq)$ defines an equivalence of categories
between the category of \'etale group schemes over $\Fq$ and the category of groups equipped
with an action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on the group.
Under this equivalence, a quasicharacter sheaf on $G$ is given by the data of:
\begin{enumerate}
 \labitem{(cs.0)}{cs.0} an indexed set of one-dimensional
  $\EE$-vector spaces $\gcs{L}_x$, as $x$ runs over
  $G(\bFq)$;

@@

 \labitem{(cs.2)}{cs.2} an indexed set of isomorphisms $\phi_{x} : \gcs{L}_{\Frob{\Fq}(x)} \to \gcs{L}_x$
  such that
  \[
   \begin{tikzcd}[row sep=40]
    \gcs{L}_{\Frob{\Fq}(x)+\Frob{\Fq}(y)} \arrow[swap]{d}{\phi_{x+y}} \arrow{rr}{\mu_{\Frob{\Fq}(x),\Frob{\Fq}(y)}}
    && \gcs{L}_{\Frob{\Fq}(x)}\otimes \gcs{L}_{\Frob{\Fq}(y)} \dar{\phi_x \tight{0}{\otimes}{0} \phi_y} \\
    \gcs{L}_{x+y} \arrow{rr}{\mu_{x,y}}
    && \gcs{L}_x \otimes\gcs{L}_y
   \end{tikzcd}
  \]
  commutes, for all $x,y\in G(\bFq)$.
\end{enumerate}
Under this equivalence, a morphism $\alpha : \cs{L} \to \cs{L'}$ of quasicharacter sheaves on $G$ is given by 
\begin{enumerate}
 \labitem{(cs.3)}{cs.3} an indexed set $\alpha_x : \gcs{L}_x \to \gcs{L'}_x$
  of linear transformations (of one-dimensional $\EE$-vector spaces),
  as $x$ runs over $G(\bFq)$, such that
  \[
   \begin{tikzcd}[column sep=40]
    \arrow[swap]{d}{\phi_x} \gcs{L}_{\Frob{\Fq}(x)} \rar{\alpha_{\Frob{\Fq}(x)}} & \gcs{L'}_{\Frob{\Fq}(x)} \dar{\phi_x'}
    &\arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}}& \arrow[swap]{d}{\mu} \gcs{L}_{x+y} \rar{\alpha_{x+y}} & \gcs{L'}_{x+y} \dar{\mu'_{x+y}} \\
    \gcs{L}_x \rar{\alpha_x} & \gcs{L'}_x
    & {} & \gcs{L}_x\otimes\gcs{L}_y \rar{\alpha_x\otimes \alpha_y} & \gcs{L'}_x \otimes \gcs{L'}_y
   \end{tikzcd}
  \]
  both commute.
\end{enumerate}

Let us say that \cdef{global section}
of a quasicharacter sheaf $\cs{L}$ on an \'etale group scheme $G$ as a choice

\begin{lemma}\label{lemma:section}
 If $\cs{L}$ is a quasicharacter sheaf on a commutative, \'etale group scheme $G$ over $\Fq$ and if $G(\bFq)$ is finitely generated, then there is a global section $s = \{ s(x) \in \gcs{L}_x \tq x \in G(\bFq) \}$ satisfying
 \begin{equation}\label{section}
  \forall x,y \in G(\bFq), \qquad \mu_{x,y}(s(x+y)) = s(x)\otimes s(y).
 \end{equation}
\end{lemma}

@@

These sections will allow us to construct the isomorphism in the following proposition.

@@

\begin{proof}
  Suppose $\cs{L}$ is a quasicharacter sheaf on $G$.  Using Lemma~\ref{lemma:section}, we may choose a section $s = \{ s(x) \in \gcs{L}_x \tq x \in G(\bFq) \}$ satisfying \eqref{section}.  We will construct a group isomorphism between $\QCiso{G(\bFq)}$ and
  $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$ using $s$, then prove independence from the choice of section.

@@

  Although we used a global section $s$ in the definition of
  $\tau_\cs{L}$, the class of $\tau_\cs{L}$ does not depend on this
  choice. To see why, let $s_1$ and $s_2$ be two global sections of
  $\cs{L}$ and let $\tau_1$ and $\tau_2$ be the cocycles defined by $s_1$ and $s_2$.
  For each $x \in G(\bFq)$ there is a unique scalar
  $a(x) \in \EEx$ such that $s_2(x) = a(x) s_1(x)$. Then \eqref{section} and
  condition~\ref{cs.1} together imply $a(x+y) = a(x)a(y)$, so
  $a\in G(\bFq)^*$.  The coboundary $\tau_0 : \Weil{\Fq} \to G(\bFq)^*$
  defined by $\tau_0(\Frob{\Fq}) = \,^\Frob{\Fq} a \cdot a^{-1}$ satisfies $\tau_2 = \tau_0 \tau_1$,
  so $\tau_2$ and $\tau_1$ are cohomologous. Thus
  \begin{equation}\label{cohomologous}
    \obj \QC(G(\bFq)) \longrightarrow \Hh^1(\Weil{\Fq},G(\bFq)^*),
    \qquad \text{defined by} \qquad \cs{L} \mapsto [\tau_\cs{L}],
  \end{equation}
  is independent of the choice of global sections made above.

@@

With Theorem~\ref{thm:snake} on hand, we can see why neither bounded quasicharacter sheaves nor finite quasicharacter sheaves are up to the task of geometrizing characters of $G(\Fq)$ in the generality we need.  
%
Consider the case when $G$ is the discrete etale group scheme $\ZZ$. Let $\chi : G(\Fq) \to \EEx$ be the character determined by $\chi(1) = \ell$ and let $\cs{L}$ be a quasicharacter sheaf in the isomorphism class corresponding to the character $\chi : G(\Fq) \to \EEx$ under Theorem~\ref{thm:snake}. Then $\trFrob{\cs{L}}(1) = \ell$. Suppose $\cs{L}$ was a bounded quasicharacter sheaf. Then the restriction of $\cs{L}$ to the connected component $1\in G$ would be a rank-1 local system on $1$, thus an $\ell$-adic character $\rho_1 : \pi_1(1,{\bar 1}) \to \EEx$ with $\rho_1(\Frob{\Fq}) = \ell$. But $\pi_1(1,{\bar 1}) = \Gal(\bFq/\Fq) \iso {\hat \ZZ}$ is compact, while $\rho_1(\pi_1(1,{\bar 1}))$ is not compact (it contains $\ell^\ZZ$). Since the Weil sheaf $\gcs{L}$ does not descend to a local system on $G$, the  quasicharacter sheaf$\cs{L}$ is not a bounded quasicharacter sheaf.
%
We can play the same game with the character $\chi : G(\Fq) \to \EEx$ defined by $\chi(1) = 1+\ell$. In this case, the corresponding geometric character $\cs{L}$ is bounded but not finite, since $1+\ell$ is not a root of unity.

These simple examples illustrate why the categories $\QCb(G)$ and $\QCf(G)$ are inadequate to geometrize characters of $G(\Fq)$.

@@

Let $K$ be a non-Archimedean local field with residue field $\Fq$.
Note that we place no restriction on the characteristic of $K$. 
We write $\OK$ for the ring of integers of $K$ and $\pK$ for the maximal ideal in $\OK$. 
For any positive integer $d$, set $A_d\ceq \OK/\pK^d$.
Let $T$ be an algebraic torus over $K$.
We write $\Hom_{< d}(T(K),\EEx)$ for the group of characters of $T(K)$ depth less than $d$ and $\Hom_\text{qc}(T(K),\EEx)$ for the group of quasicharacters of $T(K)$.

@@

@@

A lft-N\'eron model for $T$, denoted here by $\TT$, 
is a separated, smooth, commutative group scheme locally of finite type over $\OK$ with generic fibre $T$, that enjoys the N\'eron mapping property: if
$S$ is a smooth scheme over $\OK$ then every morphism $S_\eta \to T$
over $K$ extends uniquely to a morphism $S \to \TT$. Equivalently, the
N\'eron mapping property states that the function
$\Hom_{\OK} (S,\TT) \to \Hom_K(S_\eta,T)$, given by restriction to
generic fibres, is a bijection; 
in particular, taking $S = \Spec{\OK}$, this implies $\TT(\OK) = T(K)$.

\begin{definition}
For each positive integer $d$, set $\TT_{d} \ceq \TT \times_\Spec{\OK} \Spec{A_d}$. Observe that $A_d$ is an Artin ring and $\TT_{d}$ is smooth and locally of finite type over $A_d$.
%
Let \cdef{$\GN{T}_{d} = \Gr(\TT_d)$} be the group scheme over $\Fq$ produced by applying
the Greenberg functor to $\TT_{d}$; this is the \cdef{Greenberg transform of $\TT_d$}.
\end{definition}

See \cite{Greenberg:2} and \cite[Ch. 9, \S 6]{BLR} and the first few sections of \cite{Stasinski} for background on the Greenberg functor. The fundamental properties of $\GN{T}_d$ that we need are:
 \begin{enumerate}
   \labitem{(GN.0)}{dFq} $\GN{T}_d(\Fq) = \TT(A_d)$, canonically.
  \labitem{(GN.1)}{GNd} $\GN{T}_d$ is a smooth, commutative group scheme, locally of finite type over $\Fq$, and the component group of $\GN{T}_d$ is geometrically finitely generated.
  \end{enumerate}
In fact, $\pi_0(\GN{T}_d) = \pi_0(\TT) \times_{\Spec{\OK}} \Spec{\Fq}$, and the latter is finitely generated by \cite{Xarles}.

%\input{application}
\begin{theorem}\label{thm:application}
For each positive integer $d$, the trace of Frobenius 
  \[
t_{T,d}:  \QCiso{\GN{T}_d} \to \Hom_{< d}(T(K),\EEx)
  \]
is an isomomorphism from the additive functor
\[T \mapsto \QCiso{\GN{T}}\]
to the additive functor
\[
T \mapsto \Hom_{<d}(T(K),\EEx).
\]
\end{theorem}

\begin{proof}
  By \ref{GNd}, we may apply Theorem~\ref{thm:snake} to
  $\GN{T}_d$ which provides a canonical isomorphism
  $\QCiso{\GN{T}_d} \iso \GN{T}_d(\Fq)^*$.
  By \ref{dFq}, $\GN{T}_d(\Fq) = \TT(A_d)$.
  To finish the proof we recall that the filtration of $T(K)$ used to
  define depth has the property that $\TT(A_d) \cong T(K) / T(K)_d$
  \cite[Prop 5.2]{Yu}, so characters of $\GN{T}_d(\Fq)$ 
  are in bijection with characters of $T(K)$ vanishing on $T(K)_d$.

  What do we mean by saying
\end{proof}

@@

\begin{definition}
The \emph{Greenberg transform of the N\'eron model of $T$} is the group pro-scheme over $\Fq$ given by
\begin{equation}\label{projlim}
  \cdef{\GN{T}} \ceq \varprojlim_{d\in \ZZ_{>0}} \GN{T}_{d}.
\end{equation}
Let$\QC(\GN{T})$ be the ind-category $\varinjlim_{d\in \ZZ_{>0}} \QC(\GN{T}_d)$;
we refer to objects of $\QC(\GN{T})$ as \emph{quasicharacter sheaves for $T$}.
\end{definition}

\begin{remark}
The Greenberg transform of the N\'eron model of $T$ is a commutative group scheme over $\Fq$.
\end{remark}

This gives us a natural way to geometrize all quasicharacters of $T(K)$ without reference to depth.

\begin{corollary}
Let $T$ be an algebraic torus over a non-Archimedean local field $K$.
Then
  \[
 \QCiso{\GN{T}} \iso \Hom_\text{ad}(T(K),\EEx),
  \]
  canonically.
\end{corollary}
